---
title: "Platt P-I Models"
author: "S. Wotherspoon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


The package provides self starting models to fit Platt's model for a
Photosynthesis-Irradiance (PI) curve.  The `platt0` model fits a
Platt's original model

\[
P = P_{max} (1 - \exp(-\alpha I/P_{max})) \exp(-\beta I/P_{max})
\]
while `platt0` fits a modified form of the model that includes an 
additional intercept
\[
P = P_{max} (1 - \exp(-\alpha I/P_{max})) \exp(-\beta I/P_{max}) + R
\]


## PI Example

The `PI` data set contains data for 6 Photosynthesis-Irrandaiance relationships
```{r}
library(Platt)
data(PI)
head(PI)
```


```{r}
library(ggplot2)
ggplot(PI,aes(x=I,y=P,group=Depth))+
  geom_point()+
  facet_wrap(~Depth)
```


### Individual Sample

Extract the data for just depth C
```{r}
PI.C <- subset(PI,Depth=="C")
```

To fit the modified model to just this data
```{r}
fit <- nls(P~SSPlatt(I,alpha,beta,Pmax,R),data=PI.C)
summary(fit)
```

Fitted values can be extracted with `fitted`
```{r}
fitted(fit)
```
and predictions can be made for new points with `predict`
```{r}
PI.pr <- data.frame(I=seq(0,1700,20))
PI.pr$P <- predict(fit,PI.pr)
head(PI.pr)
```

The respiration corrected \(P_{max}\) is computed with `correctedPmax`
```{r}
with(as.list(coef(fit)),
  correctedPmax(alpha,beta,Pmax))
```

These can be used to compare the data and fitted model
```{r}
library(ggplot2)
ggplot(PI.C,aes(x=I,y=P))+
  geom_line(data=PI.pr,col="dodgerblue")+
  geom_point()+
  geom_point(aes(x=I,y=fitted(fit)),col="dodgerblue")
```


Plotting the residuals versus the fitted values suggests the variance
might grow with the fitted mean. 
```{r}
plot(residuals(fit)~fitted(fit))
```

### Batch Fitting

Summaries can be produced for all depths simultaneously with `by`
```{r}
by(PI,PI$Depth,function(d) summary(nls(P~SSPlatt(I,alpha,beta,Pmax,R),data=d)))
```

Alternately, a table of the estimated coefficients can be constructed
with `dplyr` 
```{r}
library(dplyr)
PI.coef <- PI %>% 
  group_by(Depth) %>% 
  do(as.data.frame(t(coef(nls(P~SSPlatt(I,alpha,beta,Pmax,R),data=.))))) %>%
  as.data.frame

PI.coef
```


Similarly, the fits can be graphically examined 
```{r}
library(dplyr)
PI.pr <- PI %>% 
  group_by(Depth) %>% 
  do({
    d <- data.frame(I=seq(0,1700,20))
    d$P <- predict(nls(P~SSPlatt(I,alpha,beta,Pmax,R),data=.),d)
    d
  })
ggplot(PI,aes(x=I,y=P,group=Depth))+
  geom_line(data=PI.pr)+
  geom_point()+
  facet_wrap(~Depth)
```


